\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample

### 2017-10-27 document/XSGRB_sample.tex
Intro itaration
A little
A little more work on abstract, intro and conclusions.

\abstract{
In this work we present the spectra of \textit{all} $\gamma$-ray bursts (GRBs)
afterglows that have \textit{ever} been observed at the VLT/X-shooter up to the end of 2016. This
totals 137 spectroscopic observation of 118 individual bursts, of which
redshifts have been secured for 91 \% covering a redshift range from 0.059 to
$\sim$ 8.5. Based on a set of unbiasing selection criteria, the follow-up effort
has been guided by the \textit{Swift}-telescope to produce a homogeneous sample
of afterglows for 86 spectra, representing the underlying \textit{Swift} GRB population.
We here provide a public release of all the spectra, including continuum
estimates and telluric corrections. We provide an assessment of the degree of
completeness with respect to the parent GRB population, in terms of the X-ray
properties of the bursts in the sample and find that the sample presented here
can not be significantly distinguished from the \textit{Swift} sample.
For all bursts (39) for which it is possible, we provide a measurement of the
optically derived hydrogen column density, increasing the total number of
published HI column densities by $\sim$ 50 \% and characterize the distribution
compared with literature values.
We constrain the fraction of dark burst to be 27\% - 42\% \todo{Preliminary} and tentatively confirm previous
results that larger optical darkness leads to increased X-ray absorption.
%This dataset provides a unique possibility to study the ISM across cosmic time.
}

@@

The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd
generation instruments at the ESO Very Large Telescope (VLT). It is a
high-throughput echelle spectrograph covering the full spectral range from the
UV atmospheric cutoff around 300 nm to the near-infrared (near-IR) K-band 
ending around 2500 nm with three spectral
arms (UVB, VIS, NIR). It was designed very much with transient source follow-up
in mind as the fading luminosities of such sources makes in urgent to secure as
extensive coverage as possible in the shortest possible time. At the same time
the resolution was set to be in the range 4000--9000 in order to be able to get
a large useful spectral range between the many sky-background emission lines in
the red and near-IR spectral ranges. In this paper we present the results
of a dedicated effort over the years 2009 -- 2016 to use the X-shooter
spectrograph to secure spectroscopic observations of
afterglows and host galaxies of GRBs detected by {\it Swift}. 
All data resulting from the survey are publicly availble in reduced form.

@@

Our sample serves the purpose to further characterize the environments of GRBs
that was also much advanced by the previous surveys based primarily on
lower-resolution spectographs. The most striking result found is that GRBs
typically are located behind very large column densities of neutral hydrogen.
Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
towards about 10$^5$ quasars have a handful such systems with
$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified \citep[e.g., five
in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, reveal such
systems in the majority of cases \citep[][and this work]{Jakobsson2006,
	Fynbo2009,Cucchiara2015}.  With afterglow spectroscopy (throughout the electromagnetic
spectrum from X-rays to the sub-mm) we hence characterize the properties of
star-forming galaxies over cosmic history in terms of redshifts, metallicities,
molecular contents, ISM temperatures, UV-flux densities, extinction curves,
etc.  A number of independent papers have been published or submitted for
publication focusing on many of these specific issues of our sample such as
extinction curves [][see also
\cite{Fynbo2014,Heintz2017}]\citep{Japelj2015,Zafar17}, emission lines from the
underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
\ion{Mg}{ii} absorbers \citep{Christensen2017}, and escape of ionizing
radiation \citep{Tanvir17}.

>>>>

\abstract{
In this work we present the spectra of \textit{all} $\gamma$-ray bursts (GRBs)
afterglows that have \textit{ever} been observed at the VLT/X-shooter up to the end of 2016. This
totals 137 spectroscopic observation of 118 individual bursts, of which
redshifts have been secured for 91 \% covering a redshift range from 0.059 to
$\sim$ 8.5. Based on a set of unbiasing selection criteria, the follow-up effort
has been guided by the \textit{Swift}-telescope to produce a homogeneous sample
of afterglows for 86 spectra, representing the underlying \textit{Swift} GRB population.
We here provide a public release of all the spectra, including continuum
estimates and telluric corrections. We provide an assessment of the degree of
completeness with respect to the parent GRB population, in terms of the X-ray
properties of the bursts in the sample and find that the sample presented here
can not be significantly distinguished from the \textit{Swift} sample.
For all bursts (39) for which it is possible, we provide a measurement of the
optically derived hydrogen column density, increasing the total number of
published HI column densities by $\sim$ 50 \% and characterize the distribution
compared with literature values.
We constrain the fraction of dark burst to be 27\% - 42\% \todo{Preliminary} and 
tentatively confirm previous
results that larger optical darkness leads is correlated with increased X-ray absorption.
%This dataset provides a unique possibility to study the ISM across cosmic time.
}

@@

The X-shooter spectrograph \citep{Vernet2011} is the first of the 2nd
generation instruments at the ESO Very Large Telescope (VLT). It is a
high-throughput echelle spectrograph covering the full spectral range from the
UV atmospheric cutoff around 300 nm to the near-infrared (near-IR) K-band 
ending around 2500 nm with three spectral
arms (UVB, VIS, NIR). It was designed very much with transient source follow-up
in mind as the fading luminosities of such sources makes in urgent to secure as
extensive coverage as possible in the shortest possible time. At the same time
the resolution was set to be in the range 4000--9000 in order to be able to get
a large useful spectral range between the many sky-background emission lines in
the red and near-IR spectral ranges. In this paper we present the results
of a dedicated effort over the years 2009 -- 2016 to use the X-shooter
spectrograph to secure spectroscopic observations of
afterglows and host galaxies of GRBs detected by {\it Swift}. 
Together with this paper we make all data resulting from the survey publicly 
availble in reduced form.

@@

Our sample serves the purpose to further characterize the environments of GRBs
that was also much advanced by the previous surveys based primarily on
lower-resolution spectographs. The most striking result found is that GRBs
typically are located behind very large column densities of neutral hydrogen.
Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
towards about 10$^5$ quasars have a handful such systems with
$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified \citep[e.g., five
in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, reveal such
systems in the majority of cases \citep[][and this work]{Jakobsson2006,
	Fynbo2009,Cucchiara2015}.  With afterglow spectroscopy (throughout the electromagnetic
spectrum from X-rays to the sub-mm) we hence characterize the properties of
star-forming galaxies over cosmic history in terms of redshifts, metallicities,
molecular contents, ISM temperatures, UV-flux densities, extinction curves,
etc.  A number of independent papers have been published or submitted for
publication focusing on many of these specific issues of our sample such as
extinction curves \citep[][see also
\cite{Fynbo2014,Heintz2017}]{Japelj2015,Zafar17}, emission lines from the
underlying host galaxies \citep{Kruhler2015}, the frequency of intervening
\ion{Mg}{ii} absorbers \citep{Christensen2017}, and escape of ionizing
radiation \citep{Tanvir17}.
